---
title: "Assignment 7"
author: "[Isaiah Wright](https://github.com/iwrig001/csde502_winter_2022_iwrig)"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
output: 
    html_document:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue        
---

<!--these following settings control width and also make a DIV tag for black outlines for figures-->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>


```{r setup, message=FALSE}
# key setups
pacman::p_load(tidyverse)

# captions
library(captioner)
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")

# for having DIV tags
knitr::knit_hooks$set(class = function(before, options, envir) {
  if(before){
    sprintf("<div class = '%s'>", options$class)
  }else{
    "</div>"
  }
})

# chunk options
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# path to this file name
if (!interactive()) {
    fnamepath <- as.character(sys.call(1))[2]
}
```

<!-- for a wide presentation -->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<!-- a DIV for black outline, good for figures-->
<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>






# Content


## Instructions for 1

1.	Generate a new data frame from the full public Add Health data set (http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip) that consists of a subset of at least four of the original columns. Do not use any of the variables used as examples in Lesson 7.
a.	Use R to download and unzip the file. You can use curl::curl_download() to download the file and utils::unzip() to unzip the file.
b.	At least one of the variables should be able to stratify the respondents into meaningful analytic groups.
c.	The data frame should have a “label” attribute that provides a brief but informative description of the table.
d.	The columns should be formatted as factor variables with proper value labels and ordering if applicable.
e.	The columns should have informative attributes as you see fit.


## Read data in

```{r}
# unzip the file
if(file.exists("data/AHwave1_v1.dta.zip") & !file.exists("data/AHwave1_v1.dta")){
    unzip(zipfile = "data/AHwave1_v1.dta.zip", exdir = "data")
}

# read the data
AHwave1_v1_haven <- haven::read_dta(file = "AHwave1_v1.dta")

# unzip the file
if(file.exists("21600-0001-Data.dta.zip") & !file.exists("21600-0001-Data.dta")){
    unzip(zipfile = "21600-0001-Data.dta.zip", exdir = "data")
}

# read the data
AdHealth_new <- haven::read_dta(file = "21600-0001-Data.dta")
```



## Save files into a temporary directory and as an RDS type file.


```{r}
# temp dir
mytempdir <- tempdir()

#write.csv(x = AdHealth_new, file = file.path(mytempdir, "AdHealth_new.csv"), row.names = FALSE)
saveRDS(object = AdHealth_new, file = file.path(mytempdir, "AdHealth_new.RDS"))
```

```{r}
attributes(AdHealth_new$H1GI8)$labels %>%
    t() %>%
    t()

AdHealth_new$H1GI8 %>% attributes()


#Convert variable to a factor
AdHealth_new$race <- factor(AdHealth_new$H1GI8)

head(AdHealth_new$race)
```



```{r}
library(magrittr)
# extract the labels from the column attribute
race_levels <- AdHealth_new$H1GI8 %>%
    attributes() %>%
    extract2("labels") %>%
    names()

race_levels

# create the factor variable 
AdHealth_new$race <- factor(AdHealth_new$H1GI8,
    labels = race_levels
)


# factor variable description
(tab_race <- AdHealth_new %>%
    group_by(race) %>%
    summarise(n = n()))
```



I'd like to use h1gi8 which is the single-race category to stratify my respondents into meaningful groups




## Create a figure

```{r}
ggplot(data = tab_race, aes(x = race, y = n)) +
    geom_bar(stat = "identity") +
    coord_flip()
```

## Generate a table depicting the count and percent race in sample.

```{r}
AdHealth_new %>%
    group_by(race) %>%
    summarise(n = n()) %>%
    mutate(`%` = n / sum(n) * 100) %>%
    mutate(`%` = `%` %>% round(1)) %>%
    kable() 
```


```{r}
# extract the labels from the column attribute
Marriage_levels <- AdHealth_new$H1GI15 %>%
    attributes() %>%
    extract2("labels") %>%
    names()

Marriage_levels

# create the factor variable 
AdHealth_new$Marriage <- factor(AdHealth_new$H1GI15,
    labels = Marriage_levels
)


# factor variable description
(tab_Marriage <- AdHealth_new %>%
    group_by(Marriage) %>%
    summarise(n = n()))

```


I imagine it would be best to soomehow exclude the response 7 from this category as the individuals were under the age of 15.


```{r}
AdHealth_new %>%
    group_by(
        race,
        Marriage
    ) %>%
    summarise(n = n(), .groups = "drop_last") %>%
    mutate(`%` = n / sum(n) * 100) %>%
    mutate(`%` = `%` %>% round(1)) %>%
    kable() 
    
    
```


```{r}
ggplot(data = tab_Marriage, aes(x = Marriage, y = n)) +
    geom_histogram(stat = "identity") +
    coord_flip()
```


<hr>

# Source code
File is at `r fnamepath`.

## Source code for this document

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```

## Complete Rmd code

```{r comment=''}
cat(readLines(fnamepath), sep = '\n')
```
